An Accurate Model for SARS-CoV-2 Pooled RT-PCR Test Errors

### 2021-09-12 paper.tex
spellcheck and refactor

Studies focused on pooling for SARS-CoV-2 are in a consensus that
sample dilution effects \cite{DilutionHIV, GroupDilution} are not a
concern, even for pools as large as 64 individuals \cite{PoolSize30,
  Lion, DorfmanYuvalDor, DilutionCOVID, CherifReview}. Consequently,
studies assumed that the probability of a true-positive (the test's
\emph{sensitivity}) does not depend on the number of infected samples
in the pool, but rather on the existence of at least one such
sample. Thus, the probability of a positive result in a pooled test
has been (assumed) identical for a pool with one sample from an
infected individual and, e.g., five such samples. This assumption is
common in the group testing literature \cite{Kim, OptimalDorfmanPool,
  CherifReview}, as well as in more specific, SARS-CoV-2 focused studies
\cite{Simplistic1, Simplistic2}. In this study we challenge this
commonly made assumption and demostrate how using a more accurate
probabilistic model affects the estimation of false-negative rates for
Dorfman pooling.

Several consequences arise from our analysis: Previously estimated optimal Dorfman
pool size \cite{OptimalDorfmanPool} should be revisited in the case of SARS-CoV-2, as it is influenced by the prevalence of infection. The error
rates of other pooling schemes in use should also be reesetimated
\cite{BayesianDorfman, Kim}. Another consequence of our results is that studies that seek precise error models should be conducted for other pathogens as well, since the standard error model could prove flawed for other pathogens. Specifically, all assumptions should be scrutinized, including (e.g.) the independence of tests, an extremely widespread assumption in the group testing literature \cite{Kim, OptimalDorfmanPool} which dates back to Dorfman's original paper \cite{DorfmanOriginal}. Finally, new methods should be developed in order to take advantage of our (and other) more realistic models.

@@

\section*{Discussion}
In this study, we developed a more realistic and novel probabilistic
model for the outcomes of pooled RT-PCR tests, parametrized for
SARS-CoV-2. Contrary to the common assumption, we assume, based on
data (Table~\ref{table}), that multiple infected individuals increase
the likelihood of a positive pooled test result. A direct consequence
of our model is that low values of infection prevalence increase the
false-negative rates of Dorfman pooling. Importantly, our results
remain qualitatively similar under varying parameter values, in the
observed ranges \cite{KitComparison,EstimatingRatesKucrika,
  EstimatingRatesLourenco, InterpretingCOVID19Test} (Figure
\ref{fig1}). Therefore, our results give rise to a conflict: low
infection prevalence leads to high efficiency of Dorfman pooling
\cite{DorfmanYuvalDor},but also increases false-negative rates.

>>>>

Studies focused on pooling for SARS-CoV-2 are in a consensus that
sample dilution effects \cite{DilutionHIV, GroupDilution} are not a
concern, even for pools as large as 64 individuals \cite{PoolSize30,
  Lion, DorfmanYuvalDor, DilutionCOVID, CherifReview}. Consequently,
studies assumed that the probability of a true-positive (the test's
\emph{sensitivity}) does not depend on the number of infected samples
in the pool, but rather on the existence of at least one such
sample. Thus, the probability of a positive result in a pooled test
has been (assumed) identical for a pool with one sample from an
infected individual and, e.g., five such samples. This assumption is
common in the group testing literature \cite{Kim, OptimalDorfmanPool,
  CherifReview}, as well as in more specific, SARS-CoV-2 focused
studies \cite{Simplistic1, Simplistic2}. In this study we challenge
this commonly made assumption and demonstrate how using a more accurate
probabilistic model affects the estimation of false-negative rates for
Dorfman pooling.

Several consequences arise from our analysis: Previously estimated
optimal Dorfman pool size \cite{OptimalDorfmanPool} should be
revisited in the case of SARS-CoV-2, as it is influenced by the
prevalence of infection. The error rates of other pooling schemes in
use should also be revisited \cite{BayesianDorfman, Kim}. Another
consequence of our results is that studies that seek precise error
models should be conducted for other pathogens as well, since the
standard error model could prove flawed for other
pathogens. Specifically, all assumptions should be scrutinized,
including (e.g.) the independence of tests, an extremely widespread
assumption in the group testing literature \cite{Kim,
  OptimalDorfmanPool} which dates back to Dorfman's original paper
\cite{DorfmanOriginal}. Finally, new methods should be developed in
order to take advantage of our (and other) more realistic models.

@@

\section*{Discussion}
In this study, we developed a more realistic and novel probabilistic
model for the outcomes of pooled RT-PCR tests, parameterized for
SARS-CoV-2. Contrary to the common assumption, we assume, based on
data (Table~\ref{table}), that multiple infected individuals increase
the likelihood of a positive pooled test result. A direct consequence
of our model is that low values of infection prevalence increase the
false-negative rates of Dorfman pooling. Importantly, our results
remain qualitatively similar under varying parameter values, in the
observed ranges \cite{KitComparison,EstimatingRatesKucrika,
  EstimatingRatesLourenco, InterpretingCOVID19Test} (Figure
\ref{fig1}). Therefore, our results give rise to a conflict: low
infection prevalence leads to high efficiency of Dorfman pooling
\cite{DorfmanYuvalDor},but also increases false-negative rates.
